---
title: "Project 2"
author: "Benjamin Hendrick"
date: "February 14, 2016"
output: html_document
---

# Part A: Data Pre-processing and Model Evaluation

## Problem 1
The formula for *min-max normalization* from the text is:

$$v_{i}' = \frac{v_{i}- min_{A}}{max_{A} - min_{A}}(newMax_{A} - newMin_{A}) + newMin_{A}$$

The R function for this formula may be written as:

```{r}
minMaxNorm <- function (x,l,u){
  normalized <- (x-min(x))/(max(x)-min(x))*(u-l)+l
  return(normalized)
}
```

where `x` is the data being normalized, `l` is the new lower bound, and `u` is the new upper bound.

The formula for *z-score normalization* from the text is: 

$$v_{i}' = \frac{v_{i} - \bar{A}}{\sigma_{A}}$$

The R function for this formula may be written as:

```{r}
zScoreNorm <- function (x,type){
  if (type == 1) {
    z <- (x-mean(x))/sd(x)
    return(z)
  }
  if (type == 2) {
    (x - mean(x)) / (length(x)^-1 * sum(abs(x-mean(x))))
  }
}
```

where `x` is the data being normalized, and `type` is the type of z-score normalization being selection (1 uses standard deviation and 2 uses mean absolute deviation.)

## Problem 2
Load the data from the text.

```{r}
data <- c(200,300,400,600,1000)
```

### Part A
Normalize the data with min-max normalization with $min=0$ and $max=1$.

```{r}
minMaxNorm(data,0,1)
```

### Part B

Normalize the data with z-score normalization using standard deviation.

```{r}
zScoreNorm(data,1)
```

### Part C

Normalize the data with z-score normalization using mean absolute deviation instead of standard deviation.

```{r}
zScoreNorm(data,2)
```


The normalized data, using all three methods in parts A, B, and C, may be summarized in the following table.

| Orignal Data | Min-Max Normalization | Z-Score Normalization (s.d.) | Z-Score Normalization (m.a.d.) |
|--------------|-----------------------|------------------------------|--------------------------------|
| 200          | 0.000                 | -0.9486833                   | -1.2500000                     |
| 300          | 0.125                 | -0.6324555                   | -0.8333333                     |
| 400          | 0.250                 | -0.3162278                   | -0.4166667                     |
| 600          | 0.500                 | 0.3162278                    | 0.4166667                      |
| 1000         | 0.100                 | 1.5811388                    | 2.0833333                      |

## Problem 3


