---
title: "CS 4821 Project 5"
author: "Benjamin Hendrick"
date: "Friday, April 08, 2016"
output: html_document
---

# Question 1

## Part A

Given the database of transactions (with min-support = 2),  there are $2^{8}-1 = 255$ frequent patterns.

$$\binom{8}{1}+ \binom{8}{2}+ \ldots +\binom{8}{8}=2^{8}-1=255$$

The eight items in the frequent item set are $I1, I2, I3, I4, I5, I6, I7,$ and $I7$

## Part B

The set of frequent *closed* items is:

- $\{I1, I2, I3, I4, I5, I5, I6, I7, I8\} : 2$
- $\{I1, I2, I3, I4, I5\} : 1$
- $\{I6, I7, I8\} : 1$

## Part C

The set of frequent *max*-patterns is:

- $\{I1, I2, I3, I4, I5, I5, I6, I7, I8\} : 2$

## Part D

An example of the association rule $\{I1, I2, I3, I4, I5, I* \rightarrow I* \}$ with min-support = 2 and min-confidence = 70% is $\{I1, I2, I3, I4, I5, I6 \rightarrow I7 \}$

$I1, I2, I3, I4, I5, I6,$ and $I7$ all occur at least twice. The confidence in the association rule is $c = \frac{2}{2} = 1$.

## Part E

For the association rule $I1 \rightarrow I6$:

- Support: $s = \frac{2}{5} = 0.4$
- Confidence: $c= \frac{2}{3} = 0.66$
- Lift: $l = \frac{2/3}{3/5} = 1.11$
- Interest: $i = \frac{2/5}{(3/5)^2} = 1.11$









# Question 2

## Part A

Refer to other submitted documents. The document `APRIORI.pdf` contains the apriori. The document `FP.pdf` contains the FP-growth tree.

## Part B

Given the rule, the *strong* association rules are 

- $\{O,K\} \rightarrow \{E\} : s= 3/5, c=3/3$
- $\{O,E\} \rightarrow \{K\} : s=3/5, c=3/3$
- $\{K,E\} \rightarrow \{O\} : s =3/5, c=3/4$







# Question 3

Load the data.

```{r, include=FALSE}
library(arules)
```

```{r}
items <- read.transactions("items.txt", format = "basket", sep=",", rm.duplicates =TRUE)
```

Create the ariori.

```{r}
items.rules <- apriori(items, parameter = list(supp = 0.6, conf = 0.8))
inspect(items.rules)
```

The frequent items sets according the apriori are {M}, {O}, {K}, {E}, {Y}, {O,E}, {O,K}. This frequent item set is smaller than the one found by the manual apriori in 2.A.







# Question 4

Suppose the given contingency table

*Transactions* | Hot Dogs | Not Hot Dogs | $\sum_{row}$ 
-|----- | ----- | ----
**Hamburger** | 2000 | 500 | 2500
**Not Hamburger** | 1000 | 1500 | 2500 
$\sum_{col}$ | 3000 | 2000 | 5000

Set up the data in `R`.

```{r}
trans <- read.transactions("trans.txt", format = "basket",sep=",")
rules <- apriori(trans, parameter = list(supp = 0.25, conf = 0.5, target="rules"))
```

## Part A

Let hot dogs = $D$ and hamburgers = $B$

Given the association rule $D \rightarrow B$, we find the confidence is $c = P(B | D) = 2000/3000 \approx 67\%$ We should consider that $P(D) = 3000/5000 = 60\%$ in the first place. Because of this fact, I would not say that the association rule is good but perhaps not strong.

## Part B

Let's look at $P(D | B) = 2000/2500 = 80%$. Also, $P(B) = 2500/5000 = 50%$. Because the confidence in the association rule $B \rightarrow D$ is so much higher than the probability of $B$ I would say that the purchase between hot dogs and hamburgers is not independent. People who by hamburgers will probably buy hot dogs too. 

Because $P(B \cap D)  = 0.4 > P(B) \times P(D) = 0.5 \times 0.6 = 0.3$, the are positively correlated.

## Part C

```{r}
im <- interestMeasure(rules, measure = c("kulczynski", "cosine", "lift", "phi"), trans)
```

For the seven association rules, the following presents the Kulczynski, cosine, life, and correlation interest measures.

The `R` function `interestMeasure` does not allow `allConfidence` or `maxConfidence` for rules, such as apriori.

Association Rule | Kulczynski | Cosine | Life | Correlation
--- | --- | --- | --- | ----
1 | `r im[1,1]` | `r im[1,2]` | `r im[1,3]` | `r im[1,4]`
2 | `r im[2,1]` | `r im[2,2]` | `r im[2,3]` | `r im[2,4]`
3 | `r im[3,1]` | `r im[3,2]` | `r im[3,3]` | `r im[3,4]`
4 | `r im[4,1]` | `r im[4,2]` | `r im[4,3]` | `r im[4,4]`
5 | `r im[5,1]` | `r im[5,2]` | `r im[5,3]` | `r im[5,4]`
6 | `r im[6,1]` | `r im[6,2]` | `r im[6,3]` | `r im[6,4]`
7 | `r im[7,1]` | `r im[7,2]` | `r im[7,3]` | `r im[7,4]`


